Inferring disease architecture and predictive ability\\with LDpred2-auto

### 2022-08-18 paper/paper-infer.tex
save
describe the extended set

\title{Inferring disease architecture and predictive ability\\with LDpred2-auto}
\author{Florian Priv\'e,$^{\text{1,}*}$  and Bjarni J. Vilhj\'almsson$^{\text{1,2}}$}

@@

\noindent$^{\text{\sf 1}}$National Centre for Register-Based Research, Aarhus University, Aarhus, 8210, Denmark. \\
\noindent$^{\text{\sf 2}}$Bioinformatics Research Centre, Aarhus University, Aarhus, 8000, Denmark. \\
\noindent$^\ast$To whom correspondence should be addressed.\\

@@

@@

\subsection{Extending the set of HapMap3 variants}

@@

The UK Biobank data is available through a procedure described at \url{https://www.ukbiobank.ac.uk/using-the-resource/}. 
All code used for this paper is available at \url{https://github.com/privefl/paper-infer/tree/master/code}.
We have extensively used R packages bigstatsr and bigsnpr \cite[]{prive2017efficient} for analyzing large genetic data, packages from the future framework \cite[]{bengtsson2020unifying} for easy scheduling and parallelization of analyses on the HPC cluster, and packages from the tidyverse suite \cite[]{wickham2019welcome} for shaping and visualizing results.
The latest version of R package bigsnpr can be installed from GitHub, and a recent enough version can be installed from CRAN [TODO: VERIFY].

>>>>

\title{Inferring disease architecture and predictive ability\\with LDpred2-auto}
\author{Florian Priv\'e,$^{\text{1,}*}$ Bogdan Pasaniuc,$^{\text{2,3,4,5}}$ and Bjarni J. Vilhj\'almsson$^{\text{1,6}}$}

@@

\noindent$^{\text{\sf 1}}$National Centre for Register-based Research, Aarhus University, Aarhus, Denmark. \\
\noindent$^{\text{\sf 2}}$Bioinformatics Interdepartmental Program, University of California, Los Angeles, CA. \\
\noindent$^{\text{\sf 3}}$Department of Human Genetics, David Geffen School of Medicine, University of California, Los Angeles, CA. \\
\noindent$^{\text{\sf 4}}$Department of Pathology and Laboratory Medicine, David Geffen School of Medicine, University of California, Los Angeles, CA. \\
\noindent$^{\text{\sf 5}}$Department of Computational Medicine, David Geffen School of Medicine, University of California, Los Angeles, CA. \\
\noindent$^{\text{\sf 6}}$Bioinformatics Research Centre, Aarhus University, Aarhus, Denmark. \\
\noindent$^\ast$To whom correspondence should be addressed. \\

@@

[ALSO TALK ABOUT IMPUTATION]

@@

\subsection{Extending the set of HapMap3 variants used}

The HapMap3 variants generally provide a good coverage of the whole genome.
We recall that the set of 1,054,330 HapMap3 variants recommended to use for LDpred2 \cite[]{prive2020ldpred2} is a subset of the original set of HapMap3 variants, which does not include duplicated positions (e.g.\ multi-allelic variants), ambiguous variants (e.g.\ both 'A' and 'T' or 'C' and 'G'), and which includes SNPs only (e.g.\ no indel).
Here we propose to extend the set we have used for LDpred2 until then. 
This extension aims at making sure many genetic variants are well tagged by the extended set, which we call HapMap3+.
To design this new set, we first read all variants from the UK Biobank (UKBB) with a minor allele frequency (MAF) larger than 0.005 in the whole data (i.e.\ the one in the MFI files). 
We then compute all pairwise correlations between variants within a 1 Mb distance, restricting to squared correlations larger than 0.3, and using all unrelated UKBB individuals excluding all White British (Field 22006) to have a diverse set of populations.
Finally, we design an algorithm which aims at maximizing the tagging of previously read variants. We want to maximize $\sum_j \max_{k \in \text{HapMap3+}} r_{j,k}^2$, where $j$ spans the whole set of variants read, while $k$ spans the variants kept in the new set.
We start by including all previously used HapMap3 variants.
For the sake of simplicity, we then use a greedy approach, where we repeatedly include the variant which increases this sum most, until no variant improves it by more than 2. 
Note that we allow only non-ambiguous SNPs to be included.
This results in an extended set of 1,444,196 SNPs.

@@

The UK Biobank data is available through a procedure described at \url{https://www.ukbiobank.ac.uk/using-the-resource/}. 
All code used for this paper is available at \url{https://github.com/privefl/paper-infer/tree/master/code}.
We have extensively used R packages bigstatsr and bigsnpr \cite[]{prive2017efficient} for analyzing large genetic data, packages from the future framework \cite[]{bengtsson2020unifying} for easy scheduling and parallelization of analyses on the HPC cluster, and packages from the tidyverse suite \cite[]{wickham2019welcome} for shaping and visualizing results.
The latest version of R package bigsnpr can be installed from GitHub.%, and a recent enough version can be installed from CRAN [TODO: VERIFY].
